The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-09-24 paper.qmd
added comment about changing broom version to get pdf to knit

# load Libraries
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)
library(ggrepel)
library(urbnmapr)
library(flextable)
```

@@

### Comparing to hypothesis tests
The most obvious way to evaluate uncertainty visualizations is to compare the visualisations to statistical tests. 
If a graphic was performing signal-suppression we would expect the signal to be harder to see at higher levels of uncertainty. 
The ideal outcome is a an uncertainty visualisation where the signal is only perceivable when it would be identified by a hypothesis test. Evaluating visualisations as though they are akin to hypothesis tests is a well established concept in visualisation. 
The lineup protocol is a good example of this approach.
Lineups are a confirmatory visualisation tool that can be used to check if perceived patterns are real or merely the result of chance [@Buja2009; @Wickham2010]. 
Lineups can also be used to evaluate the effectiveness of different types of plots [@hofmann2012] and design decisions [@vanderplas2017].
The motivation behind the lineup protocol is similar to the motivations behind signal-suppression, although the lineup protocol is more explicitly tied to a specific hypothesis. 
@Patrick2023 compared standard statistical tests to the lineup-protocol, and evaluated the visualisations using the power curves that are typical for hypothesis testing. 
In a similar vein, @Kim2019 investigated how different uncertainty visualisation methods influenced user's prior beliefs, and evaluated the graphics by comparing their results to those from Bayesian inference. 
This approach is similar to the lineup protocol as it also evaluates a visualisation by comparing it to an analogous statistical calculation, however it does so using a different statistical philosophy. 
The connection with visualisation and statistical tests even extends to EDA, where @Hullman2021 commented that a visualisation can be considered a model check where the null hypothesis is "the expected" and we are looking for "the unexpected".
This is all to say that comparing visualisations to hypothesis tests is not uncharted territory.

>>>>

# load Libraries
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)
library(ggrepel)
library(urbnmapr)
library(flextable)

# Broom returns an error with Vizumap, run this if can't compile
# devtools::install_version(package = "broom", version = "1.0.4")
```

@@

### Comparing to hypothesis tests
The most obvious way to evaluate uncertainty visualizations is to compare the visualisations to statistical tests. 
If a graphic was performing signal-suppression we would expect the signal to be harder to see at higher levels of uncertainty. 
The ideal outcome is a an uncertainty visualisation where the signal is only perceivable when it would be identified by a hypothesis test. Evaluating visualisations as though they are akin to hypothesis tests is a well established concept in visualisation. 
The lineup protocol is a good example of this approach.
Lineups are a confirmatory visualisation tool that can be used to check if perceived patterns are real or merely the result of chance [@Buja2009; @Wickham2010]. 
Lineups can also be used to evaluate the effectiveness of different types of plots [@Hofmann2012] and design decisions [@vanderplas2017].
The motivation behind the lineup protocol is similar to the motivations behind signal-suppression, although the lineup protocol is more explicitly tied to a specific hypothesis. 
@Patrick2023 compared standard statistical tests to the lineup-protocol, and evaluated the visualisations using the power curves that are typical for hypothesis testing. 
In a similar vein, @Kim2019 investigated how different uncertainty visualisation methods influenced user's prior beliefs, and evaluated the graphics by comparing their results to those from Bayesian inference. 
This approach is similar to the lineup protocol as it also evaluates a visualisation by comparing it to an analogous statistical calculation, however it does so using a different statistical philosophy. 
The connection with visualisation and statistical tests even extends to EDA, where @Hullman2021 commented that a visualisation can be considered a model check where the null hypothesis is "the expected" and we are looking for "the unexpected".
This is all to say that comparing visualisations to hypothesis tests is not uncharted territory.
